---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# hildar

<!-- badges: start -->
[![R build status](https://github.com/asiripanich/hildar/workflows/R-CMD-check/badge.svg)](https://github.com/asiripanich/hildar/actions)
<!-- badges: end -->

HILDA survey data is quite large, with many columns and rows. Some waves have more than 5000 variables. The goal of this package is to provide a quick and easy way to quickly load HILDA data into your workspace by converting the datasets from `.dta`, (one of the three formats HILDA provides) to `.fst` which can be loaded much much quicker than `.dta` in R. The main function here is `hildar::fetch()`, it can be used to load multiple years of hilda in one call and also allows a subset of variables to be selected. See `?fetch` for more options.

Please note that, this package doesn't include any parts of the HILDA survey data. You must be authorised to have access to a release of HILDA survey data. 

## 1) Installation

The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("asiripanich/hildar")
```

## 2) Store HILDA files as fst

Use `setup_hildar(dir_input = "/path/to/your/hilda-stata-files", dir_output = "/path/to/save/hilda-fst-files")` to read in HILDA STATA (.dta) files and save them as `.fst` files for `hildar`. `.fst` is a binary data format that can be read very quickly, a lot faster than `.dta`.

## 3) Set `hildar.vault` R option

Let hildar knows where your `dir_output` is by setting options(hildar.vault = "dir_output"), alternatively you may provide the .dir argument in `fetch()`.

## Example

Here is how you fetch HILDA data!

```{r example}
library(hildar)
hilda_data <- fetch(years = 2001:2003, add_geography = T)
summary(hilda_data)
```
