---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# hildar

<!-- badges: start -->
[![R-CMD-check](https://github.com/asiripanich/hildar/workflows/R-CMD-check/badge.svg)](https://github.com/asiripanich/hildar/actions)
<!-- badges: end -->

![](https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcTLV9hgux5t_pK-anbgugJ8GfoMxjD6D7_nHA&usqp=CAU) 

(source: https://www.dss.gov.au/about-the-department/national-centre-for-longitudinal-data)

[HILDA survey data](https://melbourneinstitute.unimelb.edu.au/hilda) is a large panel survey with close to 20 waves (2001 - 2020). Some waves have more than 5000 variables. The goal of this package is to provide a quick and easy way to quickly query HILDA data into your R workspace. This is possible by converting the dataset of each wave from `.dta`, one of the three formats HILDA provides, to `fst` format. `[fst](https://github.com/fstpackage/fst)` is a binary format and can be read much much quicker than `.dta` in R. The main function that `hildar` provides is `hil_fetch()`. It can load multiple waves of HILDA in one call and also allows the user to select a subset of variables instead of loading everything. See `?hil_fetch` for more options.

## 1) Installation

The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("asiripanich/hildar")
```

## 2) Store HILDA files as fst

Use `hil_setup()` to read HILDA STATA (.dta) files and save them as `.fst` files. `.fst` is a binary data format that can be read very quickly, a lot faster than `.dta`.

``` r
hil_setup(dir_input = "/path/to/your/hilda-stata-files", dir_output = "/path/to/save/hilda-fst-files")
```

This will allow you to fast query the data across all waves using `hil_fetch()`. 

## 3) Set `HILDA_FST` as a global option or an R environment variable.

`hil_fetch()` requires the user to tell it where the HILDA fst files generated in the previous step are stored. You can either set this `HILDA_FST` as a global option or an R environment variable. Setting this as a persistent option for all your R sessions will make `hil_fetch()` more convinient to use. Alternatively, you can manually set it in each call using `hilda_fst_dir` argument in `hil_fetch()`. 

## Example

Here is how you can fetch HILDA data with `hildar`!

```{r}
library(hildar)

# fetch removes the HILDA year prefix from all the selected variable
# (e.g. axxx = 2001, bxxx = 2002).
hil_fetch(years = 2001:2003, add_geography = T) %>%
  summary()
```

There is a quick option to add basic demographic variables to the data, which is set to TRUE by default.

```{r}
hil_fetch(years = 2001, add_basic_vars = T) %>%
  names()
```

How about doing a quick search to find variables that you want? Use `hil_dict` which is a data.table that you can search or view HILDA variables without going to their documentation webpage. 

```{r}
head(hil_dict)
```

Let say we want to select all variables that are related to 'employment' we can easily incorporate them into the `fetch()` function.

```{r}
emp_vars <- hil_dict[grepl(pattern = "employment", label), var]
emp_vars
hilda_data <- hil_fetch(years = 2001:2003, vars = emp_vars)
dim(hilda_data)
```

Here is a summary of the dimensions of our HILDA data files.

```{r}
# the number of variables and rows in each wave
nrows_by_wave <-
  hil_fetch(years = 2001:2016, add_basic_vars = F) %>%
  .[, .(number_of_rows = .N), by = wave]

hil_dict[, unlist(wave), by = .(var, label)] %>%
  data.table::setnames("V1", "wave") %>%
  data.table::setDT() %>%
  .[, .(number_of_variables = .N), by = wave] %>%
  merge(nrows_by_wave, by = "wave")
```
