---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hildar

<!-- badges: start -->

[![R-CMD-check](https://github.com/asiripanich/hildar/workflows/R-CMD-check/badge.svg)](https://github.com/asiripanich/hildar/actions)

<!-- badges: end -->

<p align="center">
<img src="https://fbe.unimelb.edu.au/_nocache?a=3881339" title="source: https://fbe.unimelb.edu.au/_nocache?a=3881339" width="400"/>
</p>

[HILDA survey data](https://melbourneinstitute.unimelb.edu.au/hilda) is a large panel survey with close to 20 waves (2001 - 2020), and some waves have more than 5000 variables, which makes reading them into R a little challenging.

The goal of this package is to provide a quick and easy way to query HILDA data into R. 
This is possible by converting each wave of HILDA from its STATA file (`.dta`), one of the three formats HILDA provides, to `fst` format. [fst](https://github.com/fstpackage/fst) is a binary format and can be read much much quicker than `.dta` in R.

| Function name | Description                                                                                                                                                                               |
|---------------|-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| `hil_setup()` | Setup HILDA fst files for `hil_fetch() to use`.                                                                                                                                           |
| `hil_fetch()` | Fetches HILDA records based on query options.                                                                                                                                             |
| `hil_dict()`  | Shows HILDA data glossary and waves each variable is available in. This provides a convenient way to select multiple variables based on their description by passing it to `hil_fetch()`. |
| `hil_vars()`  | Returns all variables where their variable names match a regular expression. |
| `hil_labs()`  | Returns all variables where their labels match a regular expression. |
| `hil_browse()`| Opens up the HILDA data dictionary page on your default web browser. |
| `hil_crosswave_info()`| Takes a variable name and search for its cross wave information. | 
| `hil_var_details()`| Similar to `hil_crosswave_info()` but it searches for a variable's details. |


## Installation

The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("asiripanich/hildar")
```

## Setup

### 1) Store HILDA as `.fst` files

Use `hil_setup()` to read HILDA STATA (.dta) files and save them as `.fst` files. `.fst` is a binary data format that can be read very quickly, a lot faster than `.dta`.

``` r
hil_setup(dir_input = "/path/to/your/hilda-stata-files", dir_output = "/path/to/save/hilda-fst-files")
```

This will allow you to fast query the data across all waves using `hil_fetch()`.

### 2) Tell `hildar` where the HILDA `.fst` files are stored at.

`hil_fetch()` requires the user to tell it where the HILDA fst files generated in the previous step are stored. You can either set this `HILDA_FST` as a global option or an R environment variable. Setting this as a persistent option for all your R sessions will make `hil_fetch()` more convinient to use. Alternatively, you can manually set it in each call using `hilda_fst_dir` argument in `hil_fetch()`.

## Example

Once the setup is completed, you can now start fetching HILDA data with `hildar`!

```{r}
library(hildar)

# fetch removes the HILDA year prefix from all the selected variable
# (e.g. axxx = 2001, bxxx = 2002).
hil_fetch(years = 2001:2003, add_geography = T) %>%
  summary()
```

There is a quick option to add basic demographic variables to the data, which is set to TRUE by default.

```{r}
hil_fetch(years = 2001, add_basic_vars = T) %>%
  names()
```

How about doing a quick search to find variables that you want? Use `hil_dict` which is a data.table that you can search or view HILDA variables without going to their documentation webpage.

```{r}
hilda_dictionary <- hil_dict()
head(hilda_dictionary)
```

Let say we want to select all variables that are related to 'employment'. Here is how we can easily use the selected employment variables in `hil_fetch()`.

```{r}
hilda_data <- hil_fetch(years = 2001:2003, vars = hil_labs("employment"))
dim(hilda_data)
```

Or if you know the prefix of a subject area that you like to query, you can use `hil_vars(pattern)` to query all variable names that match the pattern.
For example, `hil_vars("^ff")` will get all variables in subject area 'Health' and nested area 'Heath - diet'.

```{r}
hilda_data <- hil_fetch(years = 2001:2003, vars = hil_vars("^ff"))
dim(hilda_data)
```

Here is a summary of the dimensions of our HILDA data files.

```{r}
# the number of variables and rows in each wave
nrows_by_wave <-
  hil_fetch(years = 2001:2020, add_basic_vars = F) %>%
  .[, .(number_of_rows = .N), by = wave]

hilda_dictionary[, unlist(wave), by = .(var, label)] %>%
  data.table::setnames("V1", "wave") %>%
  .[!is.na(wave), .(number_of_variables = .N), by = wave] %>%
  merge(nrows_by_wave, by = "wave")
```
